
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(readxl)
library(openxlsx)
library(stringr)
library(data.table)

```

```{css, echo=FALSE}

h1.title {
  font-family: roboto;
  color: transparent;
  font-weight: 700;
  text-align: center;
  font-size: 12px;
  padding: 0px;
  height: 0px;
}

.footer {
  #font-family: roboto;
  color: black;
  text-align: left;
  font-size: 12px;
    padding: 5px;
  font-style: italic;
}

h1 {
  #font-family: roboto;
  color: black;
  font-weight: bolder;
  text-align: center;
  font-size: 36px;
  margin-top: 0;
  margin-bottom: 30px;
  
}
h2 {
  #font-family: roboto;
  font-weight: 500;
  color: black;
  text-align: center;
  font-size: 18px;
  margin-top: 0;
  margin-bottom: 2;
}
.bignumber {
  #font-family: roboto;
  color: white;
  font-weight: 900;
  text-align: center;
  font-size: 40px;
  line-height: 65px;
  height: 65px;
  margin-top: 0;
  margin-bottom: 0;
}
body {
  color: black;
  #font-family: roboto;
  font-weight: 400;
  font-size: 18px;
}
h4 {
  font-size: 14px;
  font-style: italic;
  text-align: center;
}
  
   * {
      box-sizing: border-box;
   }
   .card {
      color: white;
      float: left;
      width: calc(25% - 10px);
      padding: 5px;
      border-radius: 10px;
      margin-left: 3px;
      margin-right: 3px;
      margin-top: 3px;
      margin-bottom: 3px;
      height: 100%;
   }

   .card p {
     #font-family: roboto;
     text-align: center;
     font-size: 14px;
     margin-bottom: 0;
   }
   .cardContainer:after {
      content: "";
      display: table;
      clear: both;
   }
   
   @media screen and (max-width: 760px) {
      .bignumber {
         font-size: 32px;
      }
     
     .card p {
         font-size: 13px;
      }


   }
   
   @media screen and (max-width: 650px) {
      .card {
         width: 100%;
      }
      h1.title {
        font-size: 22px;
      }
      
      .bignumber {
         font-size: 35px;
      }
     
     .card p {
         font-size: 14px;
      }
      
      
   }
```

```{r, echo=FALSE, warning=FALSE}


```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

today_UTC <- as.POSIXct(Sys.time()) #get todays date as UTC
today_posix <- format(today_UTC, tz="America/Los_Angeles",usetz=TRUE)
today <- as.Date(substr(as.character(today_posix), 1,10))
yesterday <- as.Date(today-1) #get yesterday's date
filename <- paste(sep="", "counted_votes_",yesterday,".csv") #put yesterdays date into filename
vbmfile <- paste(sep="", "absentee_outstanding_",yesterday, ".zip")
#voterfile <- paste(sep="","voter_registration_",yesterday,".zip")
#pretty dates
today_display <- format(today, "%A, %b. %d, %Y")

download.file("https://s3.amazonaws.com/dl.ncsbe.gov/ENRS/2024_11_05/absentee_counts_county_20241105.csv", filename) #download vote counts
download.file("https://s3.amazonaws.com/dl.ncsbe.gov/ENRS/2024_11_05/absentee_spoiled_outstanding_20241105.zip",vbmfile) #download outstanding votes

#filename <- unzip(filename)
vbmfile <- unzip(vbmfile)
ballots <- read.csv(filename) #read voter file 
vbm <- fread(vbmfile) #read vbm file

```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
  
ballots_counted_type <- ballots %>%
  group_by(county_name, abs_request_type) %>%
  summarize(count = sum(X0)) %>%
  pivot_wider(names_from = abs_request_type, values_from = count) %>%
  mutate(MAIL = CIVILIAN + MILITARY + OVERSEAS) %>%
  mutate(`EARLY VOTING` = `EARLY VOTING` + `ONE-STOP`) %>%
  select(1,3,7) %>%
  mutate(total = `EARLY VOTING` + `MAIL`) %>%
  rename(county_desc = county_name)


#ballots_counted_type[is.na(ballots_counted_type)] <- 0

```

```{r, echo=FALSE, warning=FALSE}
#number of registered voters by county
voters <- read.csv("./nc_registeredvoters.csv") %>% clean_names()
voters_county <- voters %>% select(1,22) %>% rename(county_desc = county) %>% rename(voters = total) %>% filter(county_desc != "Totals")


```

```{r, echo=FALSE, warning=FALSE}
#remove extra files that take up a lot of memory
#rm(combined_data,temp_data,txt_files,ballots,destfile,file,voterfile)
#unlink("ncvoter_files", recursive = TRUE)


```


```{r, echo=FALSE, warning=FALSE}
#percent of registered voters who have voted so far

votes_cast <- left_join(ballots_counted_type, voters_county, by ="county_desc")
votes_cast <- votes_cast %>%
  mutate(p_votes = round((total/voters)*100,1))

write.csv(votes_cast,"votes_cast.csv", row.names = FALSE)
```



```{r, echo=FALSE, warning=FALSE}

total_voters_registered = sum(voters_county$voters)

total_vbm_issued = sum(vbm$abs_sent)

total_ballots_accepted = sum(ballots_counted_type$total)
total_ballots_early = sum(ballots_counted_type$`EARLY VOTING`)
total_ballots_vbm = sum(ballots_counted_type$MAIL)
total_pct_voted_early = round(((total_ballots_accepted/total_voters_registered)*100), digits = 1)


if (total_voters_registered > 1000000) {
  total_voters_registered_display = paste(sep="", as.character(round((total_voters_registered/1000000), digits = 1)), " M")
}

if (total_vbm_issued > 1000000) {
  total_vbm_issued_display = paste(sep="", as.character(round((total_vbm_issued/1000000), digits=1)), " M")
} else {
  total_vbm_issued_display = paste(sep="", as.character(round((total_vbm_issued/1000), digits=1)), " K")
}

if (total_ballots_accepted > 1000000) {
  total_ballots_accepted_display = paste(sep="", as.character(round((total_ballots_accepted/1000000), digits=1)), " M")
} else {
  total_ballots_accepted_display = paste(sep="", as.character(round((total_ballots_accepted/1000), digits=1)), " K")
}


```

<h1>North Carolina Ballot Tracker</h1>
<h4>Last updated `r today_display` </h4>

<div class="cardContainer">
<div class="card" style="background-color:#00318b;">
<p>Voters Registered<br>
<span class="bignumber">`r total_voters_registered_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Vote-By-Mail Ballots Requested<br>
<span class="bignumber">`r total_vbm_issued_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Ballots Accepted<br>
<span class="bignumber">`r total_ballots_accepted_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>% Voted Early<br>
<span class="bignumber">`r total_pct_voted_early`%<br></span><p>
</div>
</div>

<br>
<br>

Across North Carolina, <b>`r total_pct_voted_early`%</b> of registered voters have cast their ballots in the 2024 Presidential Election.  Early voting ends Saturday, November 2. At that point, voters must wait until Tuesday, November 5 to cast their ballot at their designated precinct.

<center><h3><a href="https://abc11.com/feature/post/voter-guide-2024-north-carolina/15465861/" target="_blank">Click here for ABC11's  Voter Guide</a></h3></center>

<iframe title="Voters by County" aria-label="Map" id="datawrapper-chart-5lVI8" src="https://datawrapper.dwcdn.net/5lVI8/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="711" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>

<br>
<br>

<iframe title="Search for your county" aria-label="Dot Plot" id="datawrapper-chart-JCRtj" src="https://datawrapper.dwcdn.net/JCRtj/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="525" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>

<br>
<br>
<div class="footer">Built and designed by Maggie Green and Lindsey Feingold. Source: <a href="https://www.ncsbe.gov/results-data" target="_blank">North Carolina State Board of Elections</a>. Data usually lags by about a day.</div>


```{r, echo=FALSE, warning=FALSE}

#trying to get date from header

#x <- XLConnect::loadWorkbook("vbm-statistics.xlsm")
#df <- XLConnect::readWorksheet(x, sheet=1)

#sheet1 <- x$worksheets[[1]]

#headers <- sheet1$headerFooter

```
